---
title: "Final Project 540-Knit"
author: 'Team 6, '
date: "12/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath("C:/Users/brand/Documents/UW-MSBA_Homework/RegressionClass"))
```

Install the Necessary Packages

XLConnect

tidyverse

```#{r eval=FALSE, include=FALSE}
#install.packages("XLConnect") #installs package to read Excel spreadsheets
#install.packages("tidyverse") #installs very powerful visualization package
#install.packages("vtable")
```

Library the Necessary Packages

readxl

tidyverse

```{r echo=FALSE, include=FALSE}
library(readxl) 
library(tidyverse)
library(vtable)
library(mi)
library(dplyr)
library(car)
```


Read Census Data

```{r echo=FALSE, include=FALSE}
Census_Base <- read_excel("CY 2020 Census Income and Population Data.xlsx",
                       sheet = "CY 2020 Census Income and Popul") #puts data in Global Environment#
```

See data types

```{r Data Types}
str(Census_Base)
```

View Census data

```{r echo=FALSE, include=FALSE}
View(Census_Base)  
```

Basic Summary. Don't use this one. It's just to show the basic table. 

```{r echo=FALSE, include=FALSE}
st(Census_Base)
```

Summary with Median. USE THIS ONE.

```{r Sumbtable}
sumtable(Census_Base, 
         summ= c('N(x)',
                'countNA(x)',
                'notNA(x)',
                'min(x)',
                'pctile(x)[25]',
                'median(x)',
                'mean(x)',
                'pctile(x)[75]',
                'max(x)'))
```




```{r data renaming}
HMI <-as.numeric((Census_Base$`Household Mean Income (dollars)`))
REGION <-(Census_Base$Region)
POP <-as.numeric((Census_Base$`Total population`))
MALE<-as.numeric((Census_Base$`Male Total`))
FEMALE<-as.numeric((Census_Base$`Female Total`))
EDU <-as.numeric((Census_Base$`Bacherlors Degree or higher`))
TOT_HOUSE <-as.numeric((Census_Base$`Total Households`))
AVG_HOUSE <-as.numeric((Census_Base$`Average Household Size`))
MARRIED <-as.numeric((Census_Base$`Total Population that has been married or currently married`))
EMP_16 <-as.numeric((Census_Base$`Civilian employed population 16 years and over`))
SELF_16 <-as.numeric((Census_Base$`Self-employed in own incorporated business workers - Civilian employed population 16 years and over`))
PRV_16 <-as.numeric((Census_Base$`Private not-for-profit wage and salary workers - Civilian employed population 16 years and over`))
GOV_16 <-as.numeric((Census_Base$`Local, state, and federal government workers - Civilian employed population 16 years and over`))
SELF2_16 <-as.numeric((Census_Base$`Self-employed in own not incorporated business workers and unpaid family workers - Civilian employed population 16 years and over`))
```

Creating Scatterplots

```{r scatterplots}
ggplot(data=Census_Base, aes(x=POP, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=MALE, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=FEMALE, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=EDU, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=TOT_HOUSE, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=AVG_HOUSE, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=MARRIED, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=EMP_16, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=SELF_16, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=PRV_16, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=GOV_16, y=HMI)) + geom_point(size=2)
ggplot(data=Census_Base, aes(x=SELF2_16, y=HMI)) + geom_point(size=2)
```

To create dummy variables k-1 k=4 so use 3 dummy variables.  South has the most, so we use it as the baseline. 
```{r  dummy variable creation}
Northeast <- ifelse(Census_Base$Region == 'Northeast', 1, 0)
Midwest <- ifelse(Census_Base$Region == 'Midwest', 1, 0)
West <- ifelse(Census_Base$Region == 'West', 1, 0)
```



```{r Census Table Combined}
Census <- data.frame(HMI, POP, MALE, FEMALE, EDU, TOT_HOUSE, AVG_HOUSE, MARRIED, EMP_16, SELF_16, PRV_16, GOV_16, SELF2_16, Northeast, Midwest,West)

View(Census)
```


First Histograms

```{r histograms}
hist(HMI)
hist(POP)
hist(MALE)
hist(FEMALE)
hist(EDU)
hist(TOT_HOUSE)
hist(AVG_HOUSE)
hist(MARRIED)
hist(EMP_16)
hist(SELF_16)
hist(PRV_16)
hist(GOV_16)
hist(SELF2_16)
hist(Midwest)
hist(Northeast)
hist(West)
```

Log the data. 

```{r logs}
Census$POP <- log(Census$POP)
Census$MALE <- log(Census$MALE)
Census$FEMALE <- log(Census$FEMALE)
Census$EDU <- log(Census$EDU)
Census$TOT_HOUSE <- log(Census$TOT_HOUSE)
Census$AVG_HOUSE <- log(Census$AVG_HOUSE)
Census$MARRIED <- log(Census$MARRIED)
Census$EMP_16 <- log(Census$EMP_16)
Census$PRV_16 <- log(Census$PRV_16)
Census$GOV_16 <- log(Census$GOV_16)
Census$SELF2_16 <- log(Census$SELF2_16)
```

Histogram the log

```{r histograms log}
hist(POP)
hist(MALE)
hist(FEMALE)
hist(EDU)
hist(TOT_HOUSE)
hist(AVG_HOUSE)
hist(MARRIED)
hist(EMP_16)
hist(SELF_16)
hist(PRV_16)
hist(GOV_16)
hist(SELF2_16)
```

Geo

```{r }
ggplot(data=Census, aes(x=POP, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=MALE, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=FEMALE, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=EDU, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=TOT_HOUSE, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=AVG_HOUSE, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=MARRIED, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=EMP_16, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=SELF_16, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=PRV_16, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=GOV_16, y=HMI)) + geom_point(size=2)+geom_smooth()
ggplot(data=Census, aes(x=SELF2_16, y=HMI)) + geom_point(size=2)+geom_smooth()
```

Remove NANs and Infs/ - Infs.

```{r removal of NAN INF and -INFs}
Census[is.na(Census) | Census == "-Inf"] = NA
```

```{r Regression full table}
FullCensusFit<-lm(HMI~POP+ MALE+ FEMALE+ EDU+ TOT_HOUSE+ AVG_HOUSE+ MARRIED+ EMP_16+ SELF_16+ PRV_16+ GOV_16+ SELF2_16 + Midwest + Northeast + West, data=Census)
summary(FullCensusFit)
```
Create first subset

```{r select subset1}
Census1 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, Northeast)
Census2 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16)
Census3 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,GOV_16)
Census4 <- Census %>% select(HMI,AVG_HOUSE,EDU,SELF_16,GOV_16)
```

Fit first subset
```{r Regression Subset1}
Census1Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16, data=Census1)
Census2Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16, data=Census2)
Census3Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ GOV_16, data=Census3)
Census4Fit<-lm(HMI~ EDU+ AVG_HOUSE+ SELF_16+ GOV_16, data=Census4)
summary(Census1Fit)
summary(Census2Fit)
summary(Census3Fit)
summary(Census4Fit)
```

Create second subsets

```{r select subset2}
Census5 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, MALE)
Census6 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, FEMALE)
Census7 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, POP)
Census8 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, TOT_HOUSE)
Census9 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, MARRIED)
Census10 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, PRV_16)
Census11 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, SELF2_16)
Census12 <- Census %>% select(HMI,AVG_HOUSE,Northeast,EDU,SELF_16,GOV_16)
Census13 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,Northeast,GOV_16)
Census14 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,Northeast)
Census15 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,Northeast,SELF_16,GOV_16)
```

```{r Regression Subset2}
Census5Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 + MALE, data=Census5)
Census6Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 +FEMALE, data=Census6)
Census7Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 + POP, data=Census7)
Census8Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 + TOT_HOUSE, data=Census8)
Census9Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 + MARRIED, data=Census9)
Census10Fit<-lm(HMI~ EDU+ AVG_HOUSE+ Northeast+ SELF_16+ GOV_16 + PRV_16, data=Census10)
Census11Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ Northeast + GOV_16 + SELF2_16, data=Census11)
Census12Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ Northeast + PRV_16, data=Census12)
Census13Fit<-lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 + Northeast, data=Census13)
Census14Fit<-lm(HMI~ AVG_HOUSE+ EMP_16+ Northeast + SELF_16+ GOV_16 + SELF2_16, data=Census14)
Census15Fit<-lm(HMI~ EDU+ Northeast+ EMP_16+ SELF_16+ GOV_16 + SELF2_16, data=Census15)

summary(Census5Fit)
summary(Census6Fit)
summary(Census7Fit)
summary(Census8Fit)
summary(Census9Fit)
summary(Census10Fit)
summary(Census11Fit)
summary(Census12Fit)
summary(Census13Fit)
summary(Census14Fit)
summary(Census15Fit)
```


```{r select subset3}
Census16 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16,GOV_16, FEMALE, Northeast)
Census17 <- Census %>% select(HMI,EMP_16,EDU,SELF_16,GOV_16, FEMALE)
Census18 <- Census %>% select(HMI,AVG_HOUSE,EDU,SELF_16,GOV_16, FEMALE)
Census19 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,SELF_16,GOV_16, FEMALE)
Census20 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,GOV_16, FEMALE)
Census21 <- Census %>% select(HMI,AVG_HOUSE,EMP_16,EDU,SELF_16, FEMALE)
```

```{r Regression Subset3}
Census16Fit<-lm(HMI~ AVG_HOUSE+EMP_16+EDU+SELF_16+GOV_16+ FEMALE+ Northeast, data=Census16)
Census17Fit<-lm(HMI~ EMP_16+EDU+SELF_16+GOV_16+ FEMALE, data=Census17)
Census18Fit<-lm(HMI~ AVG_HOUSE+EDU+SELF_16+GOV_16+ FEMALE, data=Census18)
Census19Fit<-lm(HMI~ AVG_HOUSE+EMP_16+SELF_16+GOV_16+ FEMALE, data=Census19)
Census20Fit<-lm(HMI~ AVG_HOUSE+EMP_16+EDU+GOV_16+ FEMALE, data=Census20)
Census21Fit<-lm(HMI~ AVG_HOUSE+EMP_16+EDU+SELF_16+ FEMALE, data=Census21)

summary(Census16Fit)
summary(Census17Fit)
summary(Census18Fit)
summary(Census19Fit)
summary(Census20Fit)
summary(Census21Fit)
```

Creating Residuals for the top 2 fits. 
```{r Residuals for 6}

plot(Census6Fit)

HMI.stdres6 <-residuals(Census6Fit)

plot(HMI.stdres6)


plot(Census6$EDU[1:3142],HMI.stdres6, ylab = "Residuals",xlab="Education", main= "Residuals by Education")
plot(Census6$AVG_HOUSE[1:3142],HMI.stdres6, ylab = "Residuals",xlab="AVG_HOUSE", main= "Residuals by AVG_HOUSE")
plot(Census6$EMP_16[1:3142],HMI.stdres6, ylab = "Residuals",xlab="EMP_16", main= "Residuals by EMP_16")
plot(Census6$SELF_16[1:3142],HMI.stdres6, ylab = "Residuals",xlab="SELF_16", main= "Residuals by SELF_16")
plot(Census6$GOV_16[1:3142],HMI.stdres6, ylab = "Residuals",xlab="GOV_16", main= "Residuals by GOV_16")
plot(Census6$FEMALE[1:3142],HMI.stdres6, ylab = "Residuals",xlab="FEMALE", main= "Residuals by FEMALE")


#ggplot(data=Census6, aes(x=EDU[1:3142], y=HMI.stdres6))+geom_point()
#ggplot(data=Census, aes(x=AVG_HOUSE, y=HMI.stdres6))+geom_point()
#ggplot(data=Census, aes(x=EMP_16, y=HMI.stdres6))+geom_point()
#ggplot(data=Census, aes(x=SELF_16, y=HMI.stdres6))+geom_point()
#ggplot(data=Census, aes(x=GOV_16, y=HMI.stdres6))+geom_point()
#ggplot(data=Census, aes(x=FEMALE, y=HMI.stdres6))+geom_point()

```
Predictions

```{r predictions}
HMI.predict=predict(lm(HMI~ EDU+ AVG_HOUSE+ EMP_16+ SELF_16+ GOV_16 +FEMALE, data=Census6))

```

Predictions vs Actuals
```{r predictedvsActual}
ggplot(Census6Fit, aes(x = EDU, y = HMI)) +
  geom_point()+ # Points of actual values
  geom_point(aes(y = HMI.predict), shape = 1)  # Points of predicted values

ggplot(Census6Fit, aes(x = AVG_HOUSE, y = HMI)) +
  geom_point()+ # Points of actual values
  geom_point(aes(y = HMI.predict), shape = 1) # Points of predicted values

ggplot(Census6Fit, aes(x = EMP_16, y = HMI)) +
  geom_point()+ # Points of actual values
  geom_point(aes(y = HMI.predict), shape = 1) # Points of predicted values

ggplot(Census6Fit, aes(x = SELF_16, y = HMI)) +
  geom_point()+ # Points of actual values
  geom_point(aes(y = HMI.predict), shape = 2) # Points of predicted values

ggplot(Census6Fit, aes(x = GOV_16, y = HMI)) +
  geom_point()+ # Points of actual values
  geom_point(aes(y = HMI.predict), shape = 3) # Points of predicted values

ggplot(Census6Fit, aes(x = FEMALE, y = HMI)) +
  geom_point()+ # Points of actual values
  geom_point(aes(y = HMI.predict), shape = 3)

```

Check for Collinearity
```{r}
vif(Census6Fit)
```

```{r Anova One-Way}
anova(Census6Fit)
```

